### 2019-02-22 report/final-report.tex
stuff

\section{SUMMARY}

\section{INTRODUCTION}

@@

Despite their limitations, bijections and bijective synthesis form a
useful foundation on top which more general transformations between
data sets may be built.  Hence, in the second half of our of project,
we extended the regular expression-based specifications of data types
with \emph{isomorphisms} between strings, also known as
\emph{quotients}.  To synthesize transformations between sets of
strings modulo isomorphisms, we generate canonizing functions followed
by bijections (using the original bijective synthesis algorithm).  The
so-called \emph{quotient lenses}~\cite{Foster:quotient-lenses}we generate are capable of handling
``irrelevant'' differences between structures such as white space or
permutations of items in a row of data set and hence expand the set of
transformations our system can define significantly.  Finally, we explored the
synthesis of \emph{symmetric lenses}~\cite{hofmann+:lenses} , again building upon the
bijective platform that we started with.  Symmetric lenses are able to
ignore arbitrary chunks of data in A when generating target data B,
and vice-versa.  For example, a serial number specific to one data set
may be ignored when we translate to a second.  Symmetric lenses
further expand the set of allowed transformations.

We measured the effectiveness of our algorithms on a set of
benchmarks drawn from the Augeas system~\cite{augeas}---a system for
transforming and editing Linux configuration files.  Past synthesis
tools, such as FlashFill~\cite{flashfill} were unable to translate
most Augeas file formats.  However, our synthesis toolkit was able to
transform all 40 of the Augeas file formats that we analyzed.

We have written three academic papers~\cite{?}
describing and evaluating our algorithms and produced open-source our code that
implements those algorithms.

\section{METHODS, ASSUMPTIONS, AND PROCEDURES}

@@

Lens-based languages are present in variety of tools and have found mainstream
industrial use.
Boomerang~\cite{boomerang, Matching10} lenses provide
guarantees on transformations between {\em ad hoc} string document formats.
Augeas~\cite{augeas}, a popular tool that reads Linux system configuration
files, uses the \emph{get} part of a lens to transform configuration
files into a canonical tree representation that users can edit
either manually or 
programmatically.  It uses the lens's \emph{put} to merge the edited
results back into the original string format.  Other lens-based languages and
tools include 
%
GRoundTram~\cite{Hidaka2011GRoundTramAI},
%
BiFluX~\cite{DBLP:conf/ppdp/PachecoZH14}, 
%
BiYacc~\cite{DBLP:conf/staf/ZhuK0SH15},
%
Brul~\cite{DBLP:conf/etaps/ZanLKH16},
%
BiGUL \cite{DBLP:conf/pepm/KoZH16}, 
%
bidirectional variants of 
relational algebra~\cite{BohannonPierceVaughan},
spreadsheet formulas~\cite{DBLP:conf/vl/MacedoPSC14},
graph query languages~\cite{DBLP:conf/icfp/HidakaHIKMN10},
and
XML transformation languages~\cite{DBLP:conf/pepm/LiuHT07}.

@@

The inputs to the quotient lens synthesis problem, as in the bijective case,
include source and target regular expressions $S$ and $T$ and a set of example
instances.  The equivalence relations $\equiv_S$ and $\equiv_T$ are also given
as inputs to the synthesis algorithm.  As we describe below, one result of our
work is a language of \textit{quotient regular expressions} that concisely
specifies $S$ and $\equiv_S$ simultaneously.

\section{RESULTS AND DISCUSSION}

@@

We have analyzed our algorithms theoretically, proving various
soundness and completeness results.  We have also analyzed our
algorithms empirically, studying their impact on a range of real and
synthetic benchmarks.  Our full results have been published in a
series of conference papers and technical
reports~\cite{bijective,quotient,symmetric}.  In the rest of this
section, we summarize some of the key results.

@@

We first study the effectiveness of the pure bijective synthesis
algorithm by evaluating its performance on a set of 39
benchmark programs.

All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS Sierra.

@@

We have implemented QREs and the quotient lens synthesis algorithm described
above as an extension to the Boomerang
interpreter~\cite{boomerang,quotientlenses}. We have extended the existing
Optician tool to synthesize QRE lenses from QREs. We will use 
``\QOpt'' to denote our extended version of Optician, and just plain ``Optician'' to denote
the pre-QRE version of Optician. The synthesis algorithm produces Boomerang lens
values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. To
evaluate the effectiveness of QREs, QRE lenses, and QRE lens synthesis, we
conducted experiments to answer the following questions:
%
\begin{itemize}
  \item {\bf Ease of use.} Does synthesizing QRE lenses from QREs permit an
  easier development process than writing lenses by hand?
  Does synthesizing QRE lenses from QREs permit an easier development
  process than manually writing canonizers and then synthesizing lenses between
  their canonized forms?  
  \item {\bf Performance.} Is the synthesis algorithm/implementation fast enough to be
  used as part of a standard development process?
\end{itemize}
%
All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS High Sierra.

@@

To evaluate our QRE implementation, we adapted 39 lens synthesis tasks
from the benchmark suite of the original Optician system. These
benchmarks are a combination of custom benchmarks, benchmarks derived
from FlashFill~\cite{flashfill}, and benchmarks derived from
Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).  In these 39 benchmarks, 10 of
the data formats had to be modified to work with the bijectivity
constraints that Optician required. For instance, when one
representation permits whitespace where the other does not, we
modified the original version of the benchmark to allow more whitespace,
thereby restoring bijectivity (but altering the data format). With the
new QRE support, we were able to remove these alterations.  This
experience alone suggests that QREs make the lens development process
more flexible.

\paragraph*{Ease of Use}

@@

We are interested in comparing three different approaches, which vary
in the amount of synthesis used. 
In the first approach, which we call \QRESize{} for QRE Synthesis, the programmer uses
QRE lens synthesis.  She must write QRE specifications of the source and
target formats and she may give examples.
In the second approach, which we call \canonizeAndSpecSize{} for
Bijective Synthesis, the
programmer uses bijective lens synthesis \`a la Optician.
She must write canonizers by hand, along with 
regular expressions to describe the external
representations of the source and target formats. (The internal
formats can be inferred from the canonizers.) She may also
provide examples to help in the synthesis of the bijective lens.
In the third approach, which we call \LensAndSpecSize{} for No Synthesis, the
programmer writes the lens between the source and target formats
entirely by hand, including the descriptions of the source and target
formats.

@@

\paragraph*{Maintaining Competitive Performance}

@@

To assess the performance of QRE synthesis, we are interested in two different
questions. First, how does the performance of \QOpt{} compare to the performance
of Optician on benchmarks that do not require QREs? The answer to this question
tells us how much overhead we have introduced by adopting the more general
mechanism. Figure~\ref{fig:times}(a) shows that \QOpt{} was able to synthesize
all of the Optician benchmarks at a speed competitive with the old version.
There is a small amount of additional overhead introduced by QREs in calculating
the $W$ and $K$ functions, resulting in a slight decrease in performance.

Second, how much time does it take for \QOpt{} to synthesize a QRE
lens when running on a non-bijective benchmark problem?  
Figure~\ref{fig:times}(b) shows the amount of time required to infer a
lens for each of the 10 benchmark programs with nontrivial quotients.  
We find that \QOpt{} is able to synthesize all quotient lenses in
under 10~seconds, and typically finishes in under 5~seconds.

\section{CONCLUSIONS}

>>>>

\section{Summary}

\section{Introduction}

@@

Despite their limitations, bijections and bijective synthesis form a
useful foundation on top of which more general transformations between
data sets may be built.  Hence, in the second half of our of project,
we extended the regular expression-based specifications of data types
with \emph{isomorphisms} between strings, also known as
\emph{quotients}.  To synthesize transformations between sets of
strings modulo isomorphisms, we generate canonizing functions followed
by bijections (using the original bijective synthesis algorithm).  The
so-called \emph{quotient lenses}~\cite{quotientlenses} we generate are capable of handling
``irrelevant'' differences between structures such as white space or
permutations of items in a row of data set and hence expand the set of
transformations our system can define significantly.  Finally, we explored the
synthesis of \emph{symmetric lenses}~\cite{symmetric-lenses} , again building upon the
bijective platform that we started with.  Symmetric lenses are able to
ignore arbitrary chunks of data in A when generating target data B,
and vice-versa.  For example, a serial number specific to one data set
may be ignored when we translate to a second.  Symmetric lenses
further expand the set of allowed transformations.

We measured the effectiveness of our algorithms on a set of 39
benchmarks drawn from the Augeas system~\cite{augeas},  system for
transforming and editing Linux configuration files, and
FlashFill~\cite{flashfill}, an extension to Microsoft Excel for
synthesizing columns of spreadsheets.  Past synthesis
tools, such as FlashFill~\cite{flashfill} were unable to translate
most Augeas file formats.  However, our synthesis toolkit was able to
generate transforms for all 39 of the Augeas file formats that we analyzed.

We have written
papers~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}
describing and evaluating our algorithms and produced open-source 
code~\cite{GitHub} that implements those algorithms.  In this report,
we summarize some of our key methods, results and findings, focusing
on the core algorithms for bijective synthesis and quotient synthesis.

\section{Methods, Assumptions and Procedures}

@@

Lens-based languages are present in variety of tools and have found mainstream
industrial use.
Boomerang~\cite{boomerang, Matching10} lenses provide
guarantees on transformations between {\em ad hoc} string document formats.
Augeas~\cite{augeas}, a popular tool that reads Linux system configuration
files, uses the \emph{get} part of a lens to transform configuration
files into a canonical tree representation that users can edit
either manually or 
programmatically.  It uses the lens's \emph{put} to merge the edited
results back into the original string format.  Other lens-based languages and
tools include 
%
GRoundTram~\cite{Hidaka2011GRoundTramAI},
%
BiFluX~\cite{DBLP:conf/ppdp/PachecoZH14}, 
%
BiYacc~\cite{DBLP:conf/staf/ZhuK0SH15},
%
Brul~\cite{DBLP:conf/etaps/ZanLKH16},
%
bidirectional variants of 
relational algebra~\cite{BohannonPierceVaughan},
%
BiGUL \cite{DBLP:conf/pepm/KoZH16}, 
%
spreadsheet formulas~\cite{DBLP:conf/vl/MacedoPSC14},
graph query languages~\cite{DBLP:conf/icfp/HidakaHIKMN10},
and
XML transformation languages~\cite{DBLP:conf/pepm/LiuHT07}.

@@

The inputs to the quotient lens synthesis problem, as in the bijective case,
include source and target regular expressions $S$ and $T$ and a set of example
instances.  The equivalence relations $\equiv_S$ and $\equiv_T$ are also given
as inputs to the synthesis algorithm.  Definition of such equivalences
are given via a new language of \textit{quotient regular expressions}
that we have defined.  Quotient regular expressions
describe both  $S$ (or $T$) and $\equiv_S$ (or $\equiv_T$)
simulataneously in one compact notation.

\section{Results and Discussion}

@@

We have analyzed our algorithms theoretically, proving various
soundness and completeness results.  We have also analyzed our
algorithms empirically, studying their impact on a range of real and
synthetic benchmarks.  Our full results have been published in a
series of conference papers and technical
reports~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}.  In the rest of this
section, we summarize some of the key results.

@@

We first study the effectiveness of the pure bijective synthesis
algorithm by evaluating its performance on a set of 39
benchmark programs.  
%
All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS Sierra.

@@

We have extended the bijective synthesis tool to synthesize quotient lenses. We will use 
``\QOpt'' to denote our extended version of Optician, and just plain ``Optician'' to denote
the pre-quotient version of Optician. The synthesis algorithm produces Boomerang lens
values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. 
%
All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS High Sierra.

@@

We analyzed the same 39 lens synthesis tasks from the original
Optician system.
We also experimented using our tool to synthesize quotient lenses
between XML, RDF and and JSON formats using data from the data.gov
database; the data consisted of census statistics, demographic
statistics, wage comparosion data, and crime index data).
Recall that to use bijective synthesis,  we had to modify 10 of 39
benchmarks make them bijective (for instance, by normalizing whitespace)
Because quotient lenses are more expressive, we were able avoid such
modifications. This
experience alone points to the benefits that quotient lens synthesis
brings to the table.

\paragraph*{Evaluating programmer effort}

@@

We are interested in comparing three different approaches, which vary
in the amount of synthesis used. 
In the first approach, which we call \QRESize{}, the programmer uses
quotient lens synthesis.  She must write QRE specifications of the source and
target formats and she may give examples.
In the second approach, which we call \canonizeAndSpecSize{} for
Bijective Synthesis, the
programmer uses bijective lens synthesis \`a la Optician.
She must write canonizers by hand, along with 
regular expressions to describe the external
representations of the source and target formats. (The internal
formats can be inferred from the canonizers.) She may also
provide examples to help in the synthesis of the bijective lens.
In the third approach, which we call \LensAndSpecSize{} for No Synthesis, the
programmer writes the lens between the source and target formats
entirely by hand, including the descriptions of the source and target
formats.

@@

\paragraph*{Quotient Lens Performance}

@@

To assess the performance of quotient synthesis, we are interested in two different
questions. First, how does the performance of \QOpt{} compare to the performance
of Optician on benchmarks that do not require quotients? The answer to this question
tells us how much overhead we have introduced by adopting the more general
mechanism. Figure~\ref{fig:times}(a) shows that \QOpt{} was able to synthesize
all of the Optician benchmarks at a speed competitive with the old version.
There is a small amount of additional overhead introduced by QREs in calculating
equivalences, resulting in a slight decrease in performance.

Second, how much time does it take for \QOpt{} to synthesize a quotient
lens when running on a non-bijective benchmark problem?  
Figure~\ref{fig:times}(b) shows the amount of time required to infer a
lens for each of the 10 benchmark programs with nontrivial quotients.  
We find that \QOpt{} is able to synthesize all quotient lenses in
under 10~seconds, and typically finishes in under 5~seconds.

\section{Conclusions}
